Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2019-04-11 man/ms.Rmd
update text

For each of the three resilience indices studied, robust two-way ANOVAs were used to test for differences between drought events (2005 and 2012) and the two studied oak populations (northern and southern aspects). These tests were used because original and log-transformed data did not follow the assumptions of normality and homogeneity of variance [@Wilcox2012]. Robust measures of central tendency (M-estimator based on Huber's Psi) were used since they were close to the mean value in all cases [@Wilcox2012]. When running the robust ANOVA test, data were boostrapped 3000 times and trimmed automatically to control the potential influence of outliers [@Field2012; @Wilcox2012]. Post-hoc differences were assessed pairwise using a similar boostrap test. All the robust ANOVA and post-hoc tests were carried out using the `WRS2` [@Mair2017] and `rcompanion` [@Mangiafico2017] R packages. The level of significance was set to 0.05 and adjusted for multiple comparisons. All analyses were carried out in R software [@R2017] using several packages (see Appendix S7).

@@

Las sequías produjeron reducciones de crecimiento en los tres sitios. De los eventos de sequía analizados, destaca por un lado que 1995 produjo la mayor reducción de crecimiento. 
VAS POR AQUI 
For 2005 and 2012 we found a greater reduction of RWI for northern site (SJ) but weaker for southern sites (CA-High and CA-Low). Yet, the lowest pointer year (*i.e.* the lowest growth) since 1950 was 1995 in all chronologies (Figure S3).

@@

## Resilience of greenness and radial tree-growth to drought events
At long-term perspective, the strength of the response of the resilience indices was related to drought severity. The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data (1900 - 2017) (Table S3), revealed a positive relation between resilience (*Rs*) and drought severity ($r^2$ = 0.336, p = 0.048) (Figure 6). A similar pattern was found for recovery (*Rc*; $r^2$ = 0.438, p = 0.019) but negative for resistance (*Rt*, $r^2$ = 0.155, p = 0.205).

For greenness and tree-growth, resilience metrics were significantly different bewteen drought events at short-term (Table 3). We obtained significantly higher Resilience (*Rs*) and Resistance (*Rt*) values for the 2012, the most severe event, than for 2005 in both variables (Tables S1-S2; Figure 7). Recovery (*Rc*) showed contrasting pattern for EVI and tree-growth. Recovery values for greenness was higher for 2005 than for 2012 drought event (Table S1, Figure 7).

>>>>

For each of the three resilience indices studied, robust two-way ANOVAs were used to test for differences between drought events (2005 and 2012) and the two studied oak populations (northern and southern aspects). These tests were used because original and log-transformed data did not follow the assumptions of normality and homogeneity of variance [@Wilcox2012]. Robust measures of central tendency (M-estimator based on Huber's Psi) were used since they were close to the mean value in all cases [@Wilcox2012]. When running the robust ANOVA test, data were boostrapped 3000 times and trimmed automatically to control the potential influence of outliers [@Field2012; @Wilcox2012]. Post-hoc differences were assessed pairwise using a similar boostrap test. All the robust ANOVA and post-hoc tests were carried out using the `WRS2` [@Mair2017] and `rcompanion` [@Mangiafico2017] R packages. The level of significance was set to 0.05 and adjusted for multiple comparisons. All analyses were carried out in R software [@R2017] using several packages (see Appendix S8).

@@

Las sequías produjeron reducciones de crecimiento en los tres sitios. De los eventos de sequía analizados, destaca por un lado que 1995 produjo la mayor reducción de crecimiento.

VAS POR AQUI 
For 2005 and 2012 we found a greater reduction of RWI for northern site (SJ) but weaker for southern sites (CA-High and CA-Low). Yet, the lowest pointer year (*i.e.* the lowest growth) since 1950 was 1995 in all chronologies (Figure S3).

@@

## Resilience of greenness and radial tree-growth to drought events
At long-term perspective, the strength of the response of the resilience indices was related to drought severity. The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data (1900 - 2017) (Table S3), revealed a positive relation between resilience (*Rs*) and drought severity ($r^2$ = 0.336, p = 0.048) (Figure 6). A similar pattern was found for recovery (*Rc*; $r^2$ = 0.438, p = 0.019) but negative for resistance (*Rt*, $r^2$ = 0.155, p = 0.205). Trees showed the highest value of tree-growth resilience for the worst drought event in our study area (1995, see Table S3).

During the last two drought events, resilience metrics for greenness and tree-growth were significantly different bewteen drought events (Table 3). We obtained significantly higher Resilience (*Rs*) and Resistance (*Rt*) values for the 2012, the most severe event, than for 2005 in both variables (Tables S1-S2; Figure 7). Recovery (*Rc*) showed contrasting pattern for EVI and tree-growth. Recovery values for greenness was higher for 2005 than for 2012 drought event (Table S1, Figure 7).
